---
title: "Cleaning and Wrangling Data"
author: "Tonya Macedo"
format: html
---

## Set Up

```{r}
#| message: false
#| warning: false

library(tidyr)
library(dplyr)
library(readr)
```

## Read Data

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Explore Data

```{r}
colnames(catch_original)

head(catch_original)

summary(catch_original)

unique(catch_original$Region)
```

## Clean and Reshape

### Use `select()`

```{r}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

colnames(catch_data)

# adding a - before column name will deselect folder
```

### Quality Check

```{r}
glimpse(catch_data)
```

### Replace values in column using `mutate()`

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```

Finding rows with NA values

```{r}
i <- which(is.na(catch_clean$Chinook))

catch_data[i,]
```

Fixing I in Chinoook column

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(condition = Chinook == "I", 
                           true = "1",
                           false = Chinook),
         Chinook = as.numeric(Chinook))

head(catch_clean)
```

### Re shape data with `pivot_longer()` and `pivot_wider()`

```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Year, Region), 
               names_to = "Species", 
               values_to = "Catch")

head(catch_long)
```


### Rename column using `rename()`

```{r}
colnames(catch_long)
catch_long <- catch_long %>%
  rename(Catch = Catch_thousand)

```

### Demonstrate Split-Apply-Combine

```{r}
mean_region <- catch_long %>%
  group_by(Region) %>% 
  summarise(mean_catch = mean(Catch))

head(mean_region)

n_region <- catch_long %>%
  group_by(Region) %>% 
  summarise(n = n())

head(n_region)
```

### Use `filter()` to subset by rows

```{r}
sse_catch <- catch_long %>%
  filter(Region == "SSE")

unique(sse_catch$Region)
```

### Sort dataframe using `arrange()`

Arrange sorts the column in a column

### Split and Combine values in columns using `separate()` and `unite`

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAU-039"))

sites_df %>%
    separate(site, c("island", "site_number"), "-")
```

```{r}
dates_df <- data.frame(
    year = c("1930",
             "1930",
             "1930"),
    month = c("12",
              "12",
              "12"),
    day = c("14",
            "15",
            "16")
)

dates_df %>%
    unite(date, year, month, day, sep = "-")
```


#### All Together using the pipe

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), 
               names_to = "species", 
               values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))

head(mean_region)
```


