---
title: "Data Visualization"
author: "Tonya Macedo"
format: html
---

## Set Up

```{r}
#| message: false
#| warning: false
#| 
library(readr)
library(dplyr)
library(tidyr)
library(forcats) # makes working with factors easier
library(ggplot2)
library(leaflet) # interactive maps
library(DT) # interactive tables
library(scales) # scale functions for visualization
library(janitor) # expedite cleaning and exploring data
library(viridis) # colorblind friendly color palette
```


## Read data

```{r}
delta_visits_raw <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.587.1&entityid=cda8c1384af0089b506d51ad8507641f")
```

## Explore the data

```{r}
#| eval: false

## Check out column names

colnames(delta_visits_raw)

## Peak at each column and class
glimpse(delta_visits_raw)

## From when to when
range(delta_visits_raw$Date)

## Which time of day?
unique(delta_visits_raw$Time_of_Day)
```

## Getting data ready for visualization

The janitor package has a function to format all of the column names to the same type. In this case, we choose snake_case

```{r}
delta_visits <- delta_visits_raw %>% 
    janitor::clean_names()

colnames(delta_visits)
```

Next, change the dataframe from wide to long format

```{r}
visits_long <- delta_visits %>% 
    pivot_longer(cols = c(sm_boat, med_boat, lrg_boat, bank_angler, scientist, cars),
                 names_to = "visitor_type",
                 values_to = "quantity") %>%
    rename(restore_loc = eco_restore_approximate_location) %>% 
    select(-notes)

## Checking the outcome
head(visits_long)
```

## Calculate daily visits 

```{r}
daily_visits_loc <- visits_long %>%
    group_by(restore_loc, date, visitor_type) %>% 
    summarise(daily_visits = sum(quantity))
    
head(daily_visits_loc)
```

## Plotting daily visits in ggplot2

Option 1

```{r}
## Option 1 - data and mapping called in the ggplot() function
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits))+
    geom_col()


## Option 2 - data called in ggplot function; mapping called in geom
ggplot(data = daily_visits_loc) +
    geom_col(aes(x = restore_loc, y = daily_visits))


## Option 3 - data and mapping called in geom
ggplot() +
    geom_col(data = daily_visits_loc,
             aes(x = restore_loc, y = daily_visits))
```

## Changing Geoms

```{r}
daily_visits_loc %>%
    separate(date, c("year", "month", "day"), sep = "-") %>%
    filter(daily_visits < 30,
           visitor_type %in% c("sm_boat", "med_boat", "lrg_boat")) %>%
    ggplot(aes(x = visitor_type, y = daily_visits)) +
    geom_boxplot()

```

```{r}
daily_visits_loc %>%
    separate(date, c("year", "month", "day"), sep = "-") %>%
    filter(daily_visits < 30,
           visitor_type %in% c("sm_boat", "med_boat", "lrg_boat")) %>%
    ggplot(aes(x = visitor_type, y = daily_visits)) +
    geom_violin()
```

```{r}
daily_visits_loc %>%
    filter(restore_loc == "Decker Island",
           visitor_type == "med_boat") %>%
    ggplot(aes(x = date, y = daily_visits)) +
    geom_line() +
    geom_point()
```


## Customizing plots

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = "blue"))+
    geom_col()
```


Filling by visitor type

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()
```


-change the labels
-change the x, y axis
-change the theme

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during July 2017 and March 2018")+
    coord_flip()+
    theme_bw()
```


```{r}
## Useful baseline themes are
theme_minimal()
theme_light()
theme_classic()
```

Move the legend to the bottom

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    theme_bw()+
    theme(legend.position = "bottom",
          axis.ticks.y = element_blank()) ## note we mention y-axis here
```


Save the theme you like

```{r}

my_theme <- theme_bw(base_size = 16) +
    theme(legend.position = "bottom",
          axis.ticks.y = element_blank())

ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    my_theme
```


Change the scale

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20))+
    my_theme
```

Finally we are going to expand the bars all the way to the axis line. In other words, remove the gap between the bars and the vertical “x-axis” line.

```{r}
ggplot(data = daily_visits_loc,
       aes(x = restore_loc, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

## Reordering things

```{r}
daily_visits_totals <- daily_visits_loc %>% 
    group_by(restore_loc) %>%
    mutate(n = sum(daily_visits)) %>% 
    ungroup()

head(daily_visits_totals)
```

Next, we will run the code for our plot adding the fct_reorder() function.

```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, n), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

Or you can descend

```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, desc(n)), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

## Change the colors

```{r}
ggplot(data = daily_visits_totals,
       aes(x = fct_reorder(restore_loc, desc(n)), y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    scale_fill_viridis_d()+
    labs(x = "Restoration Location",
         y = "Number of Visits",
         fill = "Type of Visitor",
         title = "Total Number of Visits to Delta Restoration Areas by visitor type",
         subtitle = "Sum of all visits during study period")+
    coord_flip()+
    scale_y_continuous(breaks = seq(0,120, 20), expand = c(0,0))+
    my_theme
```

## Saving Plots

```{r}
ggsave(here::here("plots/visit_restore_site_delta.jpg"), width = 12, height = 6, units = "in")
```


## Creating multiple plots

```{r}
facet_plot <- ggplot(data = daily_visits_totals,
       aes(x = visitor_type, y = daily_visits,
           fill = visitor_type))+
    geom_col()+
    facet_wrap(~restore_loc,
               scales = "free_y",
               ncol = 5,
               nrow = 2)+
    scale_fill_viridis_d()+
    labs(x = "Type of visitor",
         y = "Number of Visits",
         title = "Total Number of Visits to Delta Restoration Areas",
         subtitle = "Sum of all visits during study period")+
    theme_bw()+
    theme(legend.position = "bottom",
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank())

facet_plot
```

