---
title: "Intro to Quarto"
author: "Tonya Macedo"
format: html
editor: visual
execute: #this will set a default for the entire document 
  warning: false
  message: false
bibliography: references.bib
---

# Level 1 header

just regular text

## Level 2 header

-   bullet
-   list \* indented \* list item

1.  number
2.  list
3.  items

More plain text. I can type *italics* by using a single asterisk before and after the word.

I can type **bold** by using two asterisks before and after the word.

I can make it ***bold and italics*** by using three asterisks before and after the word.

```{r important r code}

x <- 1 + 2

```

My important result is `r x` (this is an in-line code chunk, where x is the final result). 
`this code is formatted`. 

```{r}
#ctl + alt + i creates a new chunk of code easily. 

x <- 4 * 8

height_ft <- c(5.2, 6.0, 5.7)

ceoef <- 3.14
```

| Species   | Reproductive Strategy | Morphology |
|-----------|-----------------------|------------|
| A palmata |                       |            |
| P lobata  |                       |            |
|           |                       |            |
|           |                       |            |
|           |                       |            |

Facts @oke2020
