---
title: "Quick analysis"
author: "Tonya Macedo"
format: html
editor: visual
---

# Introduction

## About the data

The data for this analysis were downloaded from the \[Arctic Data Center\] (https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Set up

```{r}
library(readr)
library(here)
```

## Read in the data

```{r}
#| message: false
#| warning: false 
bg_chem <- read_csv(here('data', 'BGchem2008data.csv'))
```

# Analysis

## Calculate summary statistics

```{r}

#|eval: false #will include code but not results
#|echo: false #will not show code in final 


### print the column names
colnames(bg_chem)

### get the general structure
str(bg_chem)


### First twenty lines
head(bg_chem, 20)

### Get a summary of all of the columns
summary(bg_chem)


### Unique values of a column

unique(bg_chem$Date)

# Shift command c creates a comment via shortcut. 
```

Calculate the summary statistics for nitrate, nitrite, ammonia, and phosphate.

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

## Calculate the Redfield ratio

Redfield ratio is the relative proportion of nitrogen to phosphorus.

```{r}
ratio <- (nitrate + nitrite + amm) / phos
```

The Redfield ratio for this dataset is \~ `r round(ratio)`.

$$
Redfield = \frac{Nitrogen}{Phosphorus}$$ \$\$

## Plot Redfield ratio

```{r}
#| echo: false # just no code, will evaluate and include the code
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

# Conclusion
